FISTA with adaptive discretisations

### 2021-01-07 paper.tex
Update bibliography

\section{Introduction}
The Fast Iterative Shrinkage-Thresholding Algorithm (FISTA) was first proposed by \citet{Beck2009} and is now a very popular algorithm for minimising the sum of two convex functions. We will write this as the problem of computing 
\begin{equation}\label{eq: min F = fD+fC}
	u^* \in \argmin_{\varf0\in\F H}\Func0(\varf0)  \qquad\text{such that}\qquad \Func0(\varf0)\coloneqq \func1(\varf0) + \func2(\varf0),
\end{equation}
where $\func1\colon\F H\to\R$ is a convex differentiable function with $L$-Lipschitz gradient and $\func2\colon\F H\to\bar\R$ is a convex function. The iterates of the FISTA algorithm will be denoted $\varf0_n\in\F H$. \citet{Beck2009} showed that $\Func0(\varf0_n) - \Func0(\varf0^*)$ converges at the optimum rate of $n^{-2}$ when $u^*\in \F H$ is finite dimensional and \cite{Chambolle2015} extended this result to the general Hilbert space setting. Many further works have gone on to demonstrate even faster practical convergence rates for slightly modified variants of FISTA in specific scenarios \citep{some references}.

@@

\subsection{Computing the convergence bound}
For \Cref{step: step 1 stability} of \Cref{sec: recipe} we look to replicate the classical bound of the form in \eqref{eq: classical FISTA} for \Cref{alg: refining FISTA}. The proofs in this step follow the classical arguments of \citep{Beck2009,Chambolle2015} very closely.

\subsubsection{Single iterations}
We first wish to understand a single iteration of \Cref{alg: refining FISTA}. This is done through the following two lemmas.
\begin{lemma}[equivalent to {\cite[Lemma 1]{Chambolle2015}}]\label{thm: descent lemma}
	\Copy{thm: descent lemma}{
		Suppose $\nabla \func1$ is 1-Lipschitz, for any $\bar{\varf0}\in \spcf0^{n-1}$ define
		$$ \varf0 \coloneqq \argmin_{\varf0\in \spcf0^{n}} \tfrac12\norm{\varf0-\bar{\varf0}+\nabla \func1(\bar{\varf0})}^2+\func2(\varf0).$$
		Then, for all $\varf2\in\bar{(\spcf0^n)^*}\supset\F H$, we have
		$$\Func0(\varf0) + \tfrac12\norm{\varf0-\Pi_n\varf2}^2 \leq \Func0(\Pi_n\varf2) + \tfrac12\norm{\bar{\varf0}-\Pi_n\varf2}^2$$
		where $\Pi_n\colon\bar{(\spcf0^n)^*}\to\spcf0^n$ is the orthogonal projection.
	}
\end{lemma}
\noindent This is exactly the result of \citep{Chambolle2015} applied to the function $\varf0\mapsto\Func0(\Pi_n\varf0)$. Applying \Cref{thm: descent lemma} to the iterates from \Cref{alg: refining FISTA} gives a more explicit inequality.

@@

As a special case, note that if $\vars3_{\Varf0}=1$ then \Cref{thm: sufficiently fast} already gives the optimal $O(N^{-2})$ convergence rate. This is in fact a special case of \citep{Aujol2015}. If $\varf0^*\in\F H$ then it is not possible to refine `too quickly' and the following lemma is not needed.

@@

The work of \citep{Duval2017a,Poon2018} and many others characterise the support of $\varf0^*$ very precisely. In particular, the support is at most $m$ distinct points and are a subset of $\{ \vvarx0\in\Domain \st |\A^*\vvard0^*|(\vvarx0) = \mu\} $ (an equivalent statement holds for the countable case). Less formally, this can also be seen from the the gradient computations in \Cref{sec: Lasso gap and gradient}, for all $\vvarx0\in\supp(\varf0^*)$ we have
\begin{equation}
	0\in \partial \Func0(\varf0^*)(\vvarx0) = [\A^*\vvard0^*](\vvarx0) + \mu\sign(\varf0^*(\vvarx0)).
\end{equation}

@@

Several variants of FISTA are compared for these examples but the key alternative shown here is the CBP discretisation. There are many methods designed to minimise the continuous Lasso energy in \eqref{eq: Lasso energy} \citep[c.f.][]{Bredies2013,Castro2016,Boyd2017,Catala2019} however, most result in a non-convex discretised problem to solve. We have focused on CBP because it approximates $\varf0^*$ through a convex discrete optimisation problem which is asymptotically exact in the limit $\meshsize\to0$. It can even be solved efficiently with FISTA which allows for direct comparison with the uniform and adaptive mesh approaches. As explained by \citep{Ekanadham2011, Duval2017b}, the idea is that for a fixed mesh, the kernels of $\A$ are expanded to first order on each pixel and a particular first order basis is also chosen. If $\varf0^*$ has only one Dirac spike in each pixel then the zeroth order information should correspond to the mass of the spike and additional first order information should determine the location.

@@

\section{Operator norms for numerical examples}
\begin{theorem}[\Cref{thm: norm bound examples}]\label{app:thm: norm bound examples}
	\Paste{thm: norm bound examples}
\end{theorem}
\todo{A very boring theorem but needs to be somewhere to justify the numerics. I don't know if the $\Delta\ll\sigma$ case is interesting to anyone else.}
\begin{proof}[Proof case 1:]
	From \Cref{thm: norm bound L2} we have 
	\begin{equation}
		(\A\A^*)_{i,j} = \IP{\1_{\F{\Varx0}_i}}{\1_{\F{\Varx0}_j}} = |\F{\Varx0}_i\cap \F{\Varx0}_j| = \splitln{|\F{\Varx0}_i|}{i=j}{0}{i\neq j}.
	\end{equation}
	Therefore, $\A\A^*$ is a diagonal matrix and $\norm{\A\A^*}_{\ell^2\to\ell^2} = \max_j |\F{\Varx0}_j|$ completes the result.
\end{proof}

>>>>

\section{Introduction}
The Fast Iterative Shrinkage-Thresholding Algorithm (FISTA) was first proposed by \citet{Beck2009} and is now a very popular algorithm for minimising the sum of two convex functions. We will write this as the problem of computing 
\begin{equation}\label{eq: min F = fD+fC}
	u^* \in \argmin_{\varf0\in\F H}\Func0(\varf0)  \qquad\text{such that}\qquad \Func0(\varf0)\coloneqq \func1(\varf0) + \func2(\varf0),
\end{equation}
where $\func1\colon\F H\to\R$ is a convex differentiable function with $L$-Lipschitz gradient and $\func2\colon\F H\to\bar\R$ is a convex function. The iterates of the FISTA algorithm will be denoted $\varf0_n\in\F H$. \citet{Beck2009} showed that $\Func0(\varf0_n) - \Func0(\varf0^*)$ converges at the optimum rate of $n^{-2}$ when $u^*\in \F H$ is finite dimensional and \cite{Chambolle2015} extended this result to the general Hilbert space setting. Many further works have gone on to demonstrate even faster practical convergence rates for slightly modified variants of FISTA \citep{Tao2016,Liang2017,Alamo2019}.

@@

\subsection{Computing the convergence bound}
For \Cref{step: step 1 stability} of \Cref{sec: recipe} we look to replicate the classical bound of the form in \eqref{eq: classical FISTA} for \Cref{alg: refining FISTA}. The proofs in this step follow the classical arguments of \citet{Beck2009,Chambolle2015} very closely.

\subsubsection{Single iterations}
We first wish to understand a single iteration of \Cref{alg: refining FISTA}. This is done through the following two lemmas.
\begin{lemma}[equivalent to {\cite[Lemma 1]{Chambolle2015}}]\label{thm: descent lemma}
	\Copy{thm: descent lemma}{
		Suppose $\nabla \func1$ is 1-Lipschitz, for any $\bar{\varf0}\in \spcf0^{n-1}$ define
		$$ \varf0 \coloneqq \argmin_{\varf0\in \spcf0^{n}} \tfrac12\norm{\varf0-\bar{\varf0}+\nabla \func1(\bar{\varf0})}^2+\func2(\varf0).$$
		Then, for all $\varf2\in\bar{(\spcf0^n)^*}\supset\F H$, we have
		$$\Func0(\varf0) + \tfrac12\norm{\varf0-\Pi_n\varf2}^2 \leq \Func0(\Pi_n\varf2) + \tfrac12\norm{\bar{\varf0}-\Pi_n\varf2}^2$$
		where $\Pi_n\colon\bar{(\spcf0^n)^*}\to\spcf0^n$ is the orthogonal projection.
	}
\end{lemma}
\noindent This is exactly the result of \citet{Chambolle2015} applied to the function $\varf0\mapsto\Func0(\Pi_n\varf0)$. Applying \Cref{thm: descent lemma} to the iterates from \Cref{alg: refining FISTA} gives a more explicit inequality.

@@

As a special case, note that if $\vars3_{\Varf0}=1$ then \Cref{thm: sufficiently fast} already gives the optimal $O(N^{-2})$ convergence rate. This is in fact a special case of \citet{Aujol2015}. If $\varf0^*\in\F H$ then it is not possible to refine `too quickly' and the following lemma is not needed.

@@

The work of \citet{Duval2017a,Poon2018} and many others characterise the support of $\varf0^*$ very precisely. In particular, the support is at most $m$ distinct points and are a subset of $\{ \vvarx0\in\Domain \st |\A^*\vvard0^*|(\vvarx0) = \mu\} $ (an equivalent statement holds for the countable case). Less formally, this can also be seen from the the gradient computations in \Cref{sec: Lasso gap and gradient}, for all $\vvarx0\in\supp(\varf0^*)$ we have
\begin{equation}
	0\in \partial \Func0(\varf0^*)(\vvarx0) = [\A^*\vvard0^*](\vvarx0) + \mu\sign(\varf0^*(\vvarx0)).
\end{equation}

@@

Several variants of FISTA are compared for these examples but the key alternative shown here is the CBP discretisation. There are many methods designed to minimise the continuous Lasso energy in \eqref{eq: Lasso energy} \citep[c.f.][]{Bredies2013,Castro2016,Boyd2017,Catala2019} however, most result in a non-convex discretised problem to solve. We have focused on CBP because it approximates $\varf0^*$ through a convex discrete optimisation problem which is asymptotically exact in the limit $\meshsize\to0$. It can even be solved efficiently with FISTA which allows for direct comparison with the uniform and adaptive mesh approaches. As explained by \citet{Ekanadham2011, Duval2017b}, the idea is that for a fixed mesh, the kernels of $\A$ are expanded to first order on each pixel and a particular first order basis is also chosen. If $\varf0^*$ has only one Dirac spike in each pixel then the zeroth order information should correspond to the mass of the spike and additional first order information should determine the location.

@@

\section{Operator norms for numerical examples}
\begin{theorem}[\Cref{thm: norm bound examples}]\label{app:thm: norm bound examples}
	\Paste{thm: norm bound examples}
\end{theorem}
\begin{proof}[Proof case 1:]
	From \Cref{thm: norm bound L2} we have 
	\begin{equation}
		(\A\A^*)_{i,j} = \IP{\1_{\F{\Varx0}_i}}{\1_{\F{\Varx0}_j}} = |\F{\Varx0}_i\cap \F{\Varx0}_j| = \splitln{|\F{\Varx0}_i|}{i=j}{0}{i\neq j}.
	\end{equation}
	Therefore, $\A\A^*$ is a diagonal matrix and $\norm{\A\A^*}_{\ell^2\to\ell^2} = \max_j |\F{\Varx0}_j|$ completes the result.
\end{proof}
